### 2018-11-09 slides/slides.Rmd
add template for xaringan slides

---
title: "draft slides"
author: "Andreas Beger, Predictive Heuristics"
date: "`r format(Sys.Date(), '%e %b %Y')`"
output: xaringan::moon_reader
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<!-- https://apreshill.github.io/data-vis-labs-2018/slides/06-slides_xaringan.html#13 -->

# First slide

---

# Second slides